<!--- R Base Graphics Plotting -->

```{r, echo=FALSE}
opts_chunk$set(fig.path="figure/baseplotting_")
```

### Important parameters ###
* **pch**: plotting symbol
* **lty**: line type
* **lwd**: line width
* **col**: plotting color
* **las**: orientation of axis labels 
* **bg**: background color
* **mar**: margin size
* **oma**: outer margin size
* **mfrow**: number of plots per row, column. Plots filled in row-wise.
* **mfcol**: number of plots per row, column. Plots filled in column-wise.

### Import functions ###
* **plot**: make scatterplot
* **lines**: add lines to plot
* **points**: add points to plot
* **text**: add text
* **title**: add title to axes or plot
* **mtext**: add margin text
* **axis**: add axis tick/labels

### Example 1 ###

```{r example1}
 x <- rnorm(100)
 y <- rnorm(100)
 plot(x, y, pch=21, mar=c(4,4,2,2),col='red',bg='black', xlim=c(-3,3), ylim=c(-3,3))
 fit <- lm(y ~ x)
 abline(fit, lwd = 3, col = "blue")
 title('My plot')
 text(-2, -2, 'My Label')
 legend("topleft", legend="Data", pch=21, pt.bg='black', col='red')
```

### Example 2 ###
To use multiple plots per page, give a `mfrow` argument to plot. Then the plots will be filled from left to right across the rows. (Specifying `mfcol` will fill in plots from top to bottom down the 
columns. )

```{r example2}
z <- rpois(100,2)
plot.new()
par(mfrow= c(2,2))
par(mar = c(3, 3, 2, 2))
plot(x, y, pch = 20, main="plot 1")
plot(x, z, pch = 19, main="plot 2")
plot(y, z, pch = 1, main="plot 3")
```

### Example 3 ###

The `points` function is useful for adding points to a plot for different groups.

```{r example3}
plot.new()
x <- rnorm(100)
y <- x + rnorm(100)
g <- gl(2, 50, labels = c("Male", "Female"))
str(g)
plot(x,y, type='n') # Draws no points
points(x[g == "Male"], y[g == "Male"], col="blue", pch=1)
points(x[g == "Female"], y[g == "Female"], col="red", pch=19)
legend("topleft", c("Male", "Female"), col=c("blue", "red"), pch=c(1,19))
```


